## network visualization
# reference 
  https://github.com/kateto/R-igraph-Network-Workshop-NetSciX
  https://github.com/jdwilson4/Network-Analysis-I
  dataset: enron email dataset

## example data
# set the working directory to the folder containing the files:
```{r}
setwd("D:\\SSS\\MSC\\dst\\git\\git\\Network-Analysis-I\\Data")
```

# reading data
```{r}
pol.data<-read.table("polblogs.txt",header=T)
```

# convert data to matrix

```{r}
pol.edge<-as.matrix(pol.data)+1
```


# convet network edge data to 'igraph' objects

# general igraph

```{r}
library(igraph)
pol.igraph<-graph.edgelist(pol.edge)
```

# undirected igraph

```{r}
pol.undirected.graph<-graph.edgelist(pol.edge,directed = F)
```

# adjacency matrix from igraph object

```{r}
pol.adjacency<-as_adj(pol.igraph)
```

# adjacency matrix visualization

```{r}
library(Matrix)
image(pol.adjacency)
```

# create network from edge list

```{r}
library(network)
pol.network<-network(pol.edge)
pol.x<-plot(pol.network,main=paste("political blog network"),usearrows=T,edge.col="grey50",vertex.col="yellow")
```

# color vertices according to known labels&plot

```{r}
pol.labels<-append(rep(4,758),rep(2,732),after = 758)
```

# color the lavels according to political party

```{r}
plot(pol.network,main="political blog network",usearrows=T,edge.col="grey50",coord=pol.x,vertex.col=pol.labels)
```



## enron email dataset 

# reading data
```{r}
setwd("D:\\SSS\\MSC\\dst\\git\\git\\Network-Analysis-I\\Data")
node.data<-read.table("email-Enron.txt",header=T)
```

# convert data to matrix

```{r}
node.edge<-as.matrix(node.data)+1
```

# convert network data to 'igraph' objects

# general igraph(directed)

```{r}
library(igraph)
node.igraph<-graph.edgelist(node.edge)
```

# undirected igraph

```{r}
node.un.graph<-graph.edgelist(node.edge,directed = F)
```

# adjacency matrix from igraph object

```{r}
node.adja<-as_adj(node.igraph)
```

# adjacency matrix visualization

```{r}
library(Matrix)
image(node.adja)
```

# create network from edge list

```{r}
library(network)
node.network<-network(node.edge)
node.x<-plot(node.network,main = paste("email network"),usearrows=T,edge.col="blue",vertex.col="red")
```

# color vertices according to known labels&plot

```{r}
node.labels<-append(rep(4,18352),rep(2,18300),after = 18352)
```

# color the lavels according to email

```{r}
plot(node.network,main="email network",usearrows=T,edge.col="purple",coord=node.x,vertex.col=node.labels)
```








